---
title: "Chimpanzee Example"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
fig_caption: yes
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup,include=FALSE}
knitr::opts_chunk$set(echo=TRUE,message = FALSE, error=TRUE, warning = FALSE , include = TRUE)

#install_github("rmcelreath/rethinking") #Install rethinking packages to obtain the data
library(tidyverse)
library(causact)
library(greta)
library(bayesplot)
library(rethinking)
library(dplyr)
library(ggplot2)
library(stringr)


```


We will work with the data Chimpanzee from statistical rethinking book we  want to build a model that predicts if the Chimpanzee (Actor) will pull to the left

$L_{i} \sim$Bernoulli$(p_{i})$

logit$(p_{i})=\alpha+\alpha_{Actor}+\alpha_{Block}+(\beta_{p}+\beta_{PC}C_{i})P_{i}$

$alpha_{Actor}\sim$ Normal $(0,\sigma_{Actor})$

$alpha_{Block}\sim$ Normal $(0,\sigma_{Block})$

$beta_{PC} \sim$ Normal (0,10)

$\beta_{p} \sim$ Normal (0,10)

$\alpha \sim$ Normal (0,10)
  
$\sigma_{Actor} \sim$ HalfCauchy (0,1)
  
$\sigma_{Block} \sim$ HalfCauchy (0,1)

Where $C$ represents the Condition partner absent (0), partner present (1), and P that the option was in the left.
 
```{r red_model, fig.keep='all', fig.cap = "Directed Acyclic Graph (DAG) obtained using Causact that represents the model",fig.width=12, fig.height=8}

data(chimpanzees)
d = chimpanzees
d = d%>%
  mutate(observation = row_number())

graph = dag_create()%>% # create the Directed acyclic graph
  dag_node("Pulled left", "L", #create a node
           rhs = bernoulli(p), #give a distribution to the node
           data = d$pulled_left)%>% #include data in the node
  dag_node("Probability","p",
           rhs = 1/(1+exp(-(a+b))),
           child = "L") %>% # Indicate the childrens of the node
  dag_node("a","a",
           rhs = alpha+alpha_Actor+alpha_block,
           child = "p")%>%
  dag_node("b","b",
           rhs = (beta_p+(beta_pc*C))*x,
           child = "p")%>%
  dag_node("Interception","alpha",
           rhs = normal(0,10),
           child = "a")%>%
  dag_node("Coeficient for the block","alpha_block",
           rhs = normal(0,sigma_b),
           child = "a")%>%
  dag_node("deviation for block","sigma_b",
           rhs = cauchy(0,1,truncation = c(0,Inf)),
           child = "alpha_block")%>%
  dag_node("Coeficient for actor","alpha_Actor",
           rhs = normal(0,sigma_actor),
           child = "a")%>%
  dag_node("Deviarion for actors","sigma_actor",
           rhs = cauchy(0,1,truncation = c(0,Inf)),
           child = "alpha_Actor")%>%
  dag_node("Coeficient for option was on left","beta_p",
           rhs = normal(0,10),
           child = "b")%>%
  dag_node("Beta for option was on left and condition","beta_pc",
           rhs = normal(0,10),
           child = "b")%>%
  dag_node("Condition","C",
           data = d$condition,
           child = "b")%>%
  dag_node("option was on left","x",
           data = d$prosoc_left,
           child = "b")%>%
  dag_plate("Actor","aa",
            data = d$actor,
            nodeLabels = c("alpha_Actor"))%>%
  dag_plate("Block","bb",
            data = d$block,
            nodeLabels = c("alpha_block")) #Indicate numbes of blocks

graph %>% dag_render() #Plot the DAG
graph %>% dag_greta(mcmc=TRUE) #generate greta code and run MCMC

```


The previous lines show the greta code generated by Causact
```{r prob, fig.cap = "Posterior means and confidence interval", fig.align='center', fig.width=6, fig.height=4}
 

tidyDrawsDF%>% dagp_plot()


```

```{r prob_1, fig.cap = "Posterior of Standard Deviation for the actor and block interceptions"}
 



drawsDF%>%ggplot(aes(x = sigma_b)) + 
  geom_density(color ="dodgerblue2",fill = "dodgerblue2",alpha=(0.3)) + 
  geom_density(aes(x=drawsDF$sigma_actor),color ="gold", fill = "gold",alpha=(0.3))+
  xlab("Standard Deviation")+
  ggtitle( "Posterior for the Standard Deviations")


```
Other way to plot the posterior distributions is to use dag_render from Causact package
```{r prob_3, fig.cap = "Posterior Distributions",fig.width=12, fig.height=8}
 

drawsDF %>% dagp_plot() #plot all the posterior distribution for
```